---
title: "Estructura y descripción de los metadatos de Ceiba"
author: "Marius Bottin"
date: "`r Sys.Date()`"
output: 
    bookdown::github_document2:
       number_sections: true
       toc: true
---


```{r setup}
require(RPostgreSQL)
require(dm)
require(png)
knitr::opts_chunk$set(cache=F,tidy.opts = list(width.cutoff = 70),
                     tidy = TRUE,
                     max.print=50,fig.path="./Fig/explor_ceiba_",echo=T,
                     collapse=F, echo=T)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  paste0("\n \\", "footnotesize","\n\n", x, "\n\n \\normalsize\n\n")
})
```
# Datadir en el servidor Ceiba

## Descripción
Todos los datos de Ceiba están organizados como carpetas en el datadir del servidor.
Se maneja después con el sistema Integrated Publishing Toolkit desarrollado por GBIF.

En cada carpeta (cada juego de datos), podemos encontrar:

* el archivo comprimido que contiene los archivos y los metadatos en formato DarwinCore completo.
* el archivo `eml.xml` que contiene los metadatos, y todas las versiones del archivo (con los nombres `eml-1.xml`, `eml-2.xml` etc)
* el archivo `publication.log` que contiene el historial de publicación/modificación del juego de datos
* archivos de descripción de los juegos de datos en "Rich Text Format" (rtf), tambien para cada versión publicada
* archivos de administración de datos y metadatos `resource.xml`
* una carpeta `sources` que contiene los datos (?)

## Extracción de los metadatos

En ssh, accedemos al servidor de ceiba desde la red del instituto:

```bash
ssh integracion@192.168.11.74
```

Extraemos 3 archivos:

* un archivo que tiene las direcciones de los archivos "eml.xml" y sus contenidos
* un archivo que contiene las direcciones de los archivos "resource.xml" y sus contenidos
* un catalogo de todos los archivos presentes en la carpeta de datos manejada por el ipt

Esos 2 archivos se pueden obtener con:

```bash
find /home/pem/datadir/ -name eml.xml -exec bash file_and_content.sh {} \; >file_and_content_result_eml 2> errors_find_file_and_content_eml
fimd /home/pem/datadir/ -name resource.xml  -exec bash file_and_content.sh {} \; >file_and_content_result_resource 2> errors_find_file_and_content_resource
find -type f /home/pem/datadir/ > result_find
```

Los archivos se pueden descargar desde la red del instituto, gracias al applicativo scp, que funciona a través de ssh.

# Analisis de los metadatos

```{r}
result_find <- readLines("../../../data_metadatos_catalogos/ceiba/result_find")
meta_ceiba <- readLines("../../../data_metadatos_catalogos/ceiba/file_and_content_result_eml")
```

Numero de archivos de metadatos

```{r}
sum(grepl("---file:.*---",meta_ceiba))
```

We search for the names/pathes of the files and we create the tables describing the adresses in the R object

```{r}
nameFilesAddr<-grep("---file:.*---",meta_ceiba)
addresses_xml<-data.frame(
  file=sub("^---file:(.*)---$","\\1",meta_ceiba[nameFilesAddr]),
  beg=nameFilesAddr+1,
  end=c(nameFilesAddr[2:length(nameFilesAddr)]-1,length(meta_ceiba))
)
```

Checking integrity and correspondances with the other objects

```{r}
any(duplicated(addresses_xml$file))
match(addresses_xml$file,result_find)
```
There might be a problem when there is an empty xml.
After manual checking, we realize that this problem creates an end before the beg, so to avoid these lines in the rest of the code, we just suppress these cases in the addresses_xml object:

```{r}
w_empty<-which((addresses_xml$beg[1:(nrow(addresses_xml)-1)]+1)==addresses_xml$beg[2:nrow(addresses_xml)])
addresses_xml[(w_empty-2):(w_empty+2),]

addresses_xml[addresses_xml$beg>addresses_xml$end,]
addresses_xml<-addresses_xml[addresses_xml$beg<addresses_xml$end,]
```


Separating by origin files:

```{r}
xml_files<-apply(addresses_xml,1,function(a,rl)paste(rl[a[2]:a[3]],sep="\n",collapse="\n"),rl=meta_ceiba)
names(xml_files)<-addresses_xml$file
```

Then reading the xml code:

```{r}
require(XML)
require(xml2)
xml_list<-lapply(xml_files,function(x)xmlToList(xmlParse(x)))
```

I gotta check this part because it does not make sense: this only take the large categories which are:

* dataset
* additionalMetadata
* .attrs

```r
names1<-names(xml_list[[1]])
names_all<-lapply(xml_list,names) 
sapply(names_all,function(x,y)!all(y==x),y=names1)
(names_fields<-unique(Reduce(c,names_all)))
mostComplete<-which.max(sapply(names_all,function(x,y)sum(y%in%x),y=names_fields))
  
```

Seems that part is useless as well:

```r
level1<-data.frame(
  name=names_fields,
  hasValue=F
)
A<-sapply(xml_list,names)
A_corres<-data.frame(
  lev0_nb=rep(1:length(A),sapply(A,length)),
  lev1_nb=unlist(lapply(A,function(x)1:length(x))),
  level1_match=unlist(lapply(A,function(x,y)match(x,y),y=level1$name))
)
LIST<- ISNULL <- logical(length=nrow(A_corres))
LENGTH <- DEPTH <- numeric(length=nrow(A_corres))
# NAMES <- list()
# for(i in 1:nrow(A_corres))LIST[i]<-is.list(xml_list[[A_corres[i,1]]][[A_corres[i,2]]])
# for(i in 1:nrow(A_corres))ISNULL[i]<-is.null(xml_list[[A_corres[i,1]]][[A_corres[i,2]]])
# for(i in 1:nrow(A_corres))LENGTH[i]<-length(xml_list[[A_corres[i,1]]][[A_corres[i,2]]])
# for(i in 1:nrow(A_corres))DEPTH[i]<-ldepth(xml_list[[A_corres[i,1]]][[A_corres[i,2]]])
# for(i in 1:nrow(A_corres))NAMES[[i]]<-names(xml_list[[A_corres[i,1]]][[A_corres[i,2]]])
# tapply(LENGTH,level1[A_corres[,3],"name"],table)
# table(unlist(NAMES[A_corres$level1_match==1]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==2]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==3]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==4]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==5]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==6]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==7]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==8]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==9]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==10]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==11]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==12]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==13]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==14]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==15]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==16]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==17]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==18]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==19]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==20]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==21]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==22]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==23]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==24]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==25]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==26]),useNA = "ifany")
# table(unlist(NAMES[A_corres$level1_match==27]),useNA = "ifany")
```


Algoritmo que permite navegar en toda la estructura de los xml.
Basado en 2 tablas:

1. tabla que describe la jerarquía de los campos potenciales
1. tabla que describe la presencia de los campos para los registros

Note: the following function get the elements from a recursive path

```{r}
recPathList<-function(listNavig,path)
{
  x=listNavig
  for(i in path)
    x <- x[[i]]
  return(x)
}
# example
```

Lo que hacemos primero es crear una matriz que contiene todo los paths, nivel por nivel que existen en la lista representando el XML.
Anotar: cuando el nivel siguiente no es una lista, no fila está añadida.

```{r}
require(collapse)
listStruct<-matrix(data=c(1:length(xml_list),rep(NA,length(xml_list)*(ldepth(xml_list)-1))),nrow=length(xml_list),ncol=ldepth(xml_list))
findNextLevelPaths<-function(li,pathParent,maxDepth)
{
  if(!is.list(recPathList(li,pathParent))){return(NULL)}
  LN<-length(recPathList(li,pathParent))
  return(cbind(
    matrix(data=pathParent,nrow=LN,byrow=T,ncol=length(pathParent)),
    1:LN,
    matrix(NA,nrow=LN,ncol=maxDepth-(length(pathParent)+1))
  ))
}
for(i in 2:ncol(listStruct))
{
  cat("number of parents",sum(apply(listStruct,1,function(x)length(na.omit(x)))==(i-1)),"\n")
listStruct<-rbind(listStruct,
                  Reduce(rbind,apply(listStruct[apply(listStruct,1,function(x)length(na.omit(x)))==(i-1),],1,function(p,l,ml)
  {
  p=p[!is.na(p)]
  findNextLevelPaths(l,p,ml)
  },l=xml_list,ml=ncol(listStruct),simplify = F))
)
}
```


Now let's apply a function to get transform this matrix in a name matrix.

```{r}
nameCurrentLevel<-apply(listStruct[apply(listStruct,1,function(x)length(na.omit(x)))>1,],1,function(x,li)
{
  A<-na.omit(x)
  last<-A[length(A)]
  path<-A[-length(A)]
  names(recPathList(li,path))[last]
},li=xml_list)
nameCurrentLevel[(length(nameCurrentLevel)-50):length(nameCurrentLevel)]

allNames<-matrix(NA,nrow=nrow(listStruct),ncol=ncol(listStruct)-1)
for(i in 1:nrow(listStruct))
{
  if(i%%10000==0)cat(i,"/",nrow(listStruct),"\n")
  for(j in 2:max(2,length(na.omit(listStruct[i,]))))
  {
    
    path<-listStruct[i,1:(j-1)]
    last<-listStruct[i,j]
    allNames[i,j-1]<-names(recPathList(xml_list,path))[last]
  }
}
```

Desde el nivel 2 hasta maximo, determinamos cual es el parent:

```{r}
level<-apply(listStruct,1,function(x)length(na.omit(x)))
parent<-integer(nrow(listStruct))
for(i in 2:max(level))
{
  m<-match(apply(listStruct[level==i,],1,function(x){x[which.max(which(!is.na(x)))]<-NA;return(x)},simplify = F),split(listStruct[level==(i-1),],row(listStruct[level==(i-1),])))
  parent[level==i]<-which(level==(i-1))[m]
}
```




Cuales son los hijos directos?

```{r}
directChildren<-list()
for(i in 1:length(parent))
{
  directChildren[[i]]<-which(parent==i)
}
```


Cuales son los path que contienen una lista
```{r}
contList<-apply(listStruct,1,function(x,li)
{path=na.omit(x);return(is.list(recPathList(li,path)))},li=xml_list)
```

Cuales son los casos que no contienen listas, pero son nulos:

```{r}
contNull<-apply(listStruct,1,function(x,li)
{path=na.omit(x);return(is.null(recPathList(li,path)))},li=xml_list)
noListButNull<-!contList&contNull
sum(!contList)
sum(noListButNull)
sum(!contList&!contNull)
```

```{r}
# which of listStruct is a non-null, non-list value
leaves<-which(!contList&!contNull)
# table of unique variable names
un_leaves<-unique(allNames[leaves,])
# correspondence between non-null, non-list values and unique variable names
m<-match(split(allNames[leaves,],row(allNames[leaves,])),split(un_leaves,row(un_leaves)))
# Correspondence between listStruct and unique variable names
corres_leaves<-integer(nrow(listStruct))
corres_leaves[leaves]<-m
```


```{r}
NAMES<-apply(un_leaves,1,function(x)paste(na.omit(x),collapse="."))
nb_byVar<-table(m)
names(nb_byVar)<-NAMES
nb_byVar[order(nb_byVar)]
#View(un_leaves[do.call("order",as.data.frame(un_leaves)),])
```

Para cada variable (hoja), necesitamos saber:

* cuantas veces aparece?
* En cuantos registros aparece?
* cual es el maximo de las repeticiones en un registro?
* cual es el minimo/maximo de la longitud del vector de valores?
* ejemplos de valores

```{r}

nbOccurrences<-table(corres_leaves[corres_leaves!=0])
nbReg<-tapply(listStruct[,1],corres_leaves,function(x)length(unique(x)))[-1]
nbRepMin<-apply(Reduce(rbind,tapply(factor(corres_leaves),listStruct[,1],table)),2,function(x)min(na.omit(x[x>0])))[-1]
nbRepMax<-apply(Reduce(rbind,tapply(factor(corres_leaves),listStruct[,1],table)),2,max)[-1]


rangeLN<-by(listStruct[corres_leaves!=0,],corres_leaves[corres_leaves!=0],FUN=function(tab,ls_xml)
  {
    ls_byVar<-apply(tab,1,FUN=function(x,l_x)recPathList(ls_xml,na.omit(x)),l_x=ls_xml,simplify=F)
    return(range(sapply(ls_byVar,length)))
  },ls_xml=xml_list)
len_min<-sapply(rangeLN,min)
len_max<-sapply(rangeLN,max)

subNames<-by(listStruct[corres_leaves!=0,],corres_leaves[corres_leaves!=0],FUN=function(tab,ls_xml)
  {
    subN<-unique(unlist(lapply(apply(tab,1,FUN=function(x,l_x)recPathList(ls_xml,na.omit(x)),l_x=ls_xml,simplify = F),names)))
    return(paste(subN,sep="|",collapse="|"))
  },ls_xml=xml_list)

examples<-as.list(by(listStruct,corres_leaves,FUN=function(tab,ls_xml)
  {
    ls_byVar<-apply(tab,1,FUN=function(x,l_x)recPathList(ls_xml,na.omit(x)),l_x=ls_xml)
  },ls_xml=xml_list)[-1])
isAttr<-grepl("\\.attr",NAMES)
```


Now we export a csv file to be able to analyse and name the variables:
```{r}
un_leaves<-data.frame(id=1:nrow(un_leaves),
                      un_leaves,
                      isAttr,
                      nbOccurrences,
                      nbReg,
                      nbRepMin,
                      nbRepMax,
                      len_min,
                      len_max,
                      subNames
                      )
un_leaves<-un_leaves[do.call("order",as.list(un_leaves[2:ncol(un_leaves)])),]

write.csv(un_leaves,file="../../../data_metadatos_catalogos/ceiba_un_leaves.csv")
```

# Exportation of a table with values and references (dataset and variables)

What we have got is:

* the table `un_leaves` which contains many characteristics of the different variables
* `listStruct` which contains all the paths in rows
* `corres_leaves`, which corresponds one on one to listStruct and which gives 0 when the path does not go to a "leaf" and gives the id of the table `un_leaves` when it is a leaf
* `xml_list` which contains all the info from the xml files, and might be accessed through the function `recPathList` with the path from `listStruct`
* `parent` which gives, for each of the line of `listStruct` the parent line, in `listStruct` as well 
* `addresses_xml` which contains, in the first column, the system path of the files xml, on the same order as the column 1 from `listStruct`: we can extract the name of the folder from there...

One of the difficulties we could have to extract the table of values is that some leaves (some of the .attr ones only) have more than one value on the leaves

```{r}
folder <- gsub("^.*/","",dirname(addresses_xml$file))
ref<-which(corres_leaves!=0)
res<-Reduce(rbind,lapply(ref,function(x,f,lS,cl,xl,p,ul)
  {
  val <- recPathList(xl,na.omit(lS[x,]))
  nameVal <- names(val)
  if(is.null(nameVal)){nameVal<-NA}
  return(data.frame(
    ref_struct=x,
    parent=p[x],
    folder=f[lS[x,1]],
    id_var=cl[x],
    subname=nameVal,
    value=as.character(val)
  ))},
  f=folder,
  lS=listStruct,
  cl=corres_leaves,
  xl=xml_list,
  p=parent,
  ul=un_leaves
  ))
save(res,file="../../../data_metadatos_catalogos/tab_variable.RData")
```

